---
title: "Homework 5"
author: "Yi Hang Khor"
date: "4/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(tidyverse)
library(lubridate)
library(ggplot2)
```

## Liquor Sales in Iowa

1. The complete dataset is too large. Download and import the alcohol sales data for Ames only: 
```{r}
dat <- readr::read_csv('2019_Iowa_Liquor_Sales.csv')
```

2. Data cleaning: 
  a. extract geographic latitude and longitude from the variable `Store Location`
```{r}
colnames(dat)[8] <- "Store_Location"
dat$Store_Location <- gsub("POINT \\(", "",dat$Store_Location)
dat$Store_Location <- gsub("\\)", "",dat$Store_Location) 
dat <- separate(dat, "Store_Location", into = c("Latitude", "Longitude"), sep = " (?=[^ ]+$)")
dat[16:20,8:9]
```
  b. make sure the variables are in the correct type, by converting incorrect variable types.
```{r}
dat$Latitude <- as.numeric(dat$Latitude)
dat$Longitude <- as.numeric(dat$Longitude)
dat$`Vendor Number` <-as.numeric(dat$`Vendor Number`)
str(dat)
```
  c. extract year, month and day from the variable `Date` as separate columns.
```{r}
dat <- separate(dat, "Date", into = c("Month","Day","Year"), sep = "/")
dat$Month <- as.numeric(dat$Month)
dat$Day <- as.numeric(dat$Day)
dat$Year <-as.numeric(dat$Year)
str(dat)
```
3. First overview:
  a. Plot a scatterplot of lat and long of store locations. 
```{r}
ggplot(dat,aes(x= Latitude,y= Longitude,color = `Store Name`))+geom_point() +xlab("Latitude") +ylab("Longitude") +ggtitle("Store Locations based on Latitude and Longitude") +theme(legend.position = 'none')
```
 b. Provide a visual breakdown of the liquor category (by `Category Name`). Include volume sold in the breakdown.
```{r}
liq <- dat%>%
  group_by(`Category Name`) %>% 
  summarize(total= (sum(`Volume Sold (Liters)`))) %>% 
  arrange(desc(total))
ggplot(liq[1:15,],aes(x = reorder(`Category Name`,-total),y = `total`,fill = `Category Name`)) +geom_bar(stat= 'identity') +xlab("Liquor Name")+ggtitle("Top 15 Best Selling Liquor") +ylab("Total Sales")+scale_x_discrete(labels = abbreviate) +theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

4. Find the daily sales liquor sales in Ames in 2019. Summarize the number of sales, volume of liquor sold, and amount of money spent.
```{r}
day <- dat %>%
  group_by(`Month`,`Day`) %>%
  summarize(sales = (sum (`Bottles Sold`)),volsold = (sum(`Volume Sold (Liters)`)),spent= (sum(`Sale (Dollars)`)))
ggplot(day,aes(x= Day,y = sales))+geom_point()+facet_wrap(vars(Month)) +ylab("Number of Sales")
ggplot(day,aes(x= Day,y = spent))+geom_point()+facet_wrap(vars(Month)) +ylab("Amount of Money Spent(Dollars)")
```

5. Plot volume sold by day (use a scatterplot of volume by day and facet by month). Discover trends with the help of `geom_smooth`. Describe any patterns.
```{r}
plt <- ggplot(day,aes(x= Day,y = volsold))+geom_point()+facet_wrap(vars(Month)) +ylab("Volume of Liquor Sold(Liters)") +geom_smooth(span = 0.3)
suppressWarnings(print(plt))
```

**Based on the graph, each amplitude (The highest Volume Sold) seems to occur in a weekly interval. **

6. Find the dates for ISU football home games in Fall 2019. Can you see a pattern? 

**Iowa State Home Games : 8/31/2019 9/14/2019 9/21/2019 10/5/2019 10/26/2019 11/16/2019 11/23/2019**
<br/>
**Due to the lack of data for ISU home games, hence, no pattern can be seen.**